---
title: "HW5 : Distribution"
author:
- Ebenezer Nkum^[enkum@miners.utep.edu]
- University of Texas at El Paso (UTEP)
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    latex_engine: pdflatex
    number_sections: yes
    toc: yes
    toc_depth: 4
  html_document:
    toc: yes
    toc_depth: '4'
    df_print: paged
geometry: margin=1in
fontsize: 11pt
spacing: single
header-includes:
- \usepackage{amsmath}
- \usepackage{amssymb}
- \usepackage{amsfonts}
- \usepackage{amsthm}
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhf{}
- "\\rhead{DS 6339-- Data Visualization}"
- \cfoot{\thepage}
- \usepackage{algorithm}
- \usepackage[noend]{algpseudocode}
- "\\DeclareUnicodeCharacter{2212}{\\ensuremath{-}}"
editor_options: 
  chunk_output_type: inline
---

<!-- QUESTION ONE: WHAT --> 
\noindent\rule{17.5cm}{0.8pt}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE, include=FALSE, message=FALSE, warning=FALSE}
# Load libraries
#library(lubridate)
library(tidyverse)
library(dplyr)
library(plotly)
library(hrbrthemes)
library(kableExtra)
library(knitr)
library(tinytex)
library(tibble)
library(ggrepel)
library("reshape2")   
```



# Load data
```{r Load data}
# Load data 
dat <- read.csv("serialdat.csv", header = T)
```

# Data

The new data (serialdat.csv), contains information about gene variant transcriptions. There were three replications of the variant transcriptions and a final column where the three replications were averaged. The two categorial varialbes included are the SUMOver (which is classes of genes - S#V# format). Ignore the replication number that follows the S#V# groups (of which there are 6 groups). Try to visualize the distributions according to each group. You may also include visualizations across the replications to see if the transcription process introduced error overall.


# Step 1: Inspect data, assess it for completeness, good formatting, and any errors
## Inspect data
```{r inspect data, echo=FALSE}

head(dat) ; dim(dat)

```

## Remove the replication number that follows the S#V# groups
```{r , message=FALSE}
dat0 <- dat[-43,] %>% select(-X10.x.copies) %>% 
                mutate(SUMOvar = sapply(str_split(SUMOvar,"-",),'[',1))

head(dat0)
```

# Visualizing distribution

# Overall distribution among each group using the average replications

```{r}
ggplot(dat0, aes(x= SUMOvar, y=Average.Cq, color = SUMOvar)) +
      geom_boxplot()+
     theme_classic() +
    ggtitle("Distribution of Average Cq by Gene Classes")+
     theme(legend.position = "none")
```
Here see that, the gene class S2V1 has more variability than all of the other classes. And more than half of its average replications lies below 19. The highest median average replication appears under the gene class S1V1. It has a median average replication of 24.



```{r}
dat_reshape <- melt(dat0[,-5], id = "SUMOvar")

colnames(dat_reshape) <- c("SUMOvar","variable", "Replications")
```

# Distributions by each group
```{r, fig.width=10}

ggplot(dat_reshape, aes(x=SUMOvar, y= Replications, fill = variable )) +
            geom_boxplot() +
            theme_classic() +
            ggtitle("Distribution of 3 replications by Gene Classes")+
            theme(legend.position = "none")+
            facet_wrap(~variable)

           
```
We observe that under each replication, S2V1 has more variability than rest of the gene classes. And also, S1V1 has the highest median average replications in all the three replications. 












